---
title: "Networks"
description: "Inferred transmission networks from time calibrated phylogenies"
format:
  html: 
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options: 
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
knit_print.gt <- function(x, ...) {
  stringr::str_c(
    "<div style='all:initial';>\n", 
    gt::as_raw_html(x), 
    "\n</div>"
  ) |> 
    knitr::asis_output()
    
}
registerS3method(
  "knit_print", 'gt_tbl', knit_print.gt, 
  envir = asNamespace("gt") 
)
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(ape)
library(ggtree)
library(ggraph)
library(igraph)
library(ggmap)
library(ggspatial)
library(dendextend)
library(TransPhylo)
library(pals)
library(gt)

## Custom Functions
source(here("R/utilities.R"))
source_dir(here("R"))
```


## Transmission Networks


## Serotype A   
Load time-calibrated MCC tree.  
```{r}
sero_A.tree <- read.nexus(here("local/beast/a_1/As.mcc.tre"))
```
  
Read Beast log file for the tree
```{r}
get_A_stats <- get_tracer_stats(here("local/beast/a_1/As_rev.log"))

root_age <- get_A_stats %>%
    filter(Parameter == "age.root.") %>%
    select(Median) %>%
    pull()
  
tree_mrsd <- get_A_stats %>%
    filter(Parameter == "treeModel.rootHeight") %>%
    select(Median) %>%
    pull() + root_age

sero_A.tree.p <- ptreeFromPhylo(sero_A.tree,
                                dateLastSample = tree_mrsd)
```
   
Choose prior  
Constructing a gamma distribution that reflects FMDV generation time, ballpark 5-15 days, with 7-10 days being more likely.  
```{r}
gam_params <- get_gamma_params(7, c(5, 15))

w.shape <- gam_params$shape
w.scale <- gam_params$scale

set.seed(1976)
check_gamma <- rgamma(1000, w.shape,   )
```
```{r fig.height=6, fig.width=6, warning=FALSE}
plot_density(check_gamma, max_x = 20)
```

The `dateT` parameter is the date when observation stopped.  Used 2017.5 due that being the most recent sample across the A, Asia1, and O serotypes in the study.  
```{r eval=FALSE}
tt_sero_A <- inferTTree(sero_A.tree.p,
                        mcmcIterations=10000,
                        w.shape=w.shape,
                        w.scale=w.scale,
                        dateT=2017.5)

saveRDS(tt_sero_A, here("local/assets/tt_sero_A.rds"))
```

## Serotype Asia1
```{r}
sero_Asia1.tree <- read.nexus(here("local/beast/asia1_1/asia1.mcc.tre"))
Asia1_stats <- get_tracer_stats(here("local/beast/asia1_1/Asia1s.log"))
```
```{r eval=FALSE}
tt_sero_Asia1 <- infer_ttree(sero_Asia1.tree,
                             Asia1_stats,
                             w.shape=w.shape,
                             w.scale=w.scale,
                             dateT = 2017.5)

saveRDS(tt_sero_Asia1, here("local/assets/tt_sero_Asia1.rds"))
```
  
## Serotype O
```{r}
sero_O.tree <- read.nexus(here("local/beast/o_1/Os.mcc.tre"))
O_stats <- get_tracer_stats(here("local/beast/o_1/sero_0.log"))
```
```{r eval=FALSE}
tt_sero_O <- infer_ttree(sero_O.tree,
                         O_stats,
                         w.shape=w.shape,
                         w.scale=w.scale,
                         dateT = 2017.5)

saveRDS(tt_sero_O, here("local/assets/tt_sero_O.rds"))
```  
   
## Get Networks 

```{r echo=FALSE}
tt_sero_A <- readRDS(here("local/assets/tt_sero_A.rds"))
tt_sero_Asia1 <- readRDS(here("local/assets/tt_sero_Asia1.rds"))
tt_sero_O <- readRDS(here("local/assets/tt_sero_O.rds"))
```
```{r}
A_net <- extract_transmission_network(tt_sero_A) %>%
  mutate(Serotype = "A")

Asia1_net <- extract_transmission_network(tt_sero_Asia1) %>%
  mutate(Serotype = "Asia1")

O_net <- extract_transmission_network(tt_sero_O) %>%
  mutate(Serotype = "O")

all_networks <- rbind(A_net, Asia1_net, O_net)
```
  
## Match Metadata
```{r}
sero_df <- readRDS(here("local/assets/sero_df.rds")) # from preprocessing

animal_codes <- sero_df %>%
  filter(status == "Subclinical") %>%
  mutate(host_id = label) %>%
  select(host_id, animal, farm_code, status)

all_networks <-left_join(all_networks, animal_codes, by = "host_id") %>%
  mutate(animal = as.character(animal),
         status = if_else(is.na(status), "Clinical", status),
         infected_ani = if_else(status == "Subclinical", animal, host_id))

all_networks$infector_match <- with(animal_codes,
                                    animal[match(
                                      all_networks$infector,
                                      host_id)])

all_networks <- all_networks %>%
  mutate(infector_match = as.character(infector_match),
         infector_ani = if_else(is.na(infector_match), infector, infector_match),
         infection_date = as.Date(infection_date))
```

## Set Vertices and Nodes
```{r}
edges <- data.frame(from = all_networks$infector_ani, to = all_networks$infected_ani)

g <- graph_from_data_frame(edges, directed = TRUE)

V(g)$status <- "Unknown"

V(g)$status[V(g)$name %in% all_networks$infected_ani] <- all_networks$status[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$Serotype <- "Origin"
V(g)$Serotype[V(g)$name %in% all_networks$infected_ani] <- 
  all_networks$Serotype[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$infection_date <- NA
V(g)$infection_date[V(g)$name %in% all_networks$infected_ani] <- 
  all_networks$infection_date[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$infection_date <- as.Date(V(g)$infection_date)
```
  
First attempt  
```{r fig.width=10, fig.height=10}

p <- ggraph(g, layout = 'tree') + 
  geom_edge_diagonal(start_cap = circle(2, 'mm'), end_cap = circle(2, 'mm'),
                 color = 'gray20', alpha=0.9, 
                 arrow = arrow(type = "closed", length = unit(2, "mm"))) +
  geom_node_point(aes(color = status, shape = Serotype), size = 4, alpha = 0.6) +
  scale_color_manual(values = c("Subclinical" = "orange", "Clinical" = "steelblue", "Unknown" = "gray50")) +
  #geom_node_label(aes(label = name), repel = TRUE) +
  theme_void() +
  labs(title = "Network Graph",
       subtitle = "Nodes: host_id, Edges: infector to host_id")

p
```

   
#### To be continued...   

```{r eval=FALSE, echo=FALSE}
library(gganimate)


edges <- data.frame(from = all_networks$infector_ani, to = all_networks$infected_ani)

g <- graph_from_data_frame(edges, directed = TRUE)

V(g)$status <- "Unknown"
V(g)$status[V(g)$name %in% all_networks$infected_ani] <- 
  all_networks$status[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$Serotype <- NA
V(g)$Serotype[V(g)$name %in% all_networks$infected_ani] <- 
  all_networks$Serotype[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$infection_date <- NA
V(g)$infection_date[V(g)$name %in% all_networks$infected_ani] <- 
  as.Date(all_networks$infection_date[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)])

nodes_df <- data.frame(name = V(g)$name, status = V(g)$status, Serotype = V(g)$Serotype, 
                       infection_date = as.Date(V(g)$infection_date))

nodes_df$infection_date[is.na(nodes_df$infection_date)] <- min(nodes_df$infection_date, na.rm = TRUE) - 1

p <- ggraph(g, layout = 'tree') + 
  geom_edge_link(start_cap = circle(2, 'mm'), end_cap = circle(2, 'mm'),
                 color = 'gray20', alpha=0.9, 
                 arrow = arrow(type = "closed", length = unit(2, "mm"))) +
  geom_node_point(aes(color = status, shape = Serotype), size = 5, alpha = 0.6) +
  scale_color_manual(values = c("Subclinical" = "orange", 
                                "Clinical" = "steelblue", 
                                "Unknown" = "gray50")) +
  theme_void() +
  labs(title = "Network Graph",
       subtitle = "Nodes: host_id, Edges: infector to host_id") +
  transition_time(nodes_df$infection_date) +
  labs(subtitle = 'Date: {frame_time}') +
    shadow_mark(alpha = alpha / 2, color = "gray40")

anim <- animate(p, width = 800, height = 600, 
                fps = 10, end_pause = 50, 
                renderer = gifski_renderer())

#anim_save("network_animation.gif", anim)
anim

```
```{r eval=FALSE, echo=FALSE}
nodes_df <- data.frame(name = V(g)$name, 
                       status = V(g)$status, 
                       Serotype = V(g)$Serotype, 
                       infection_date = V(g)$infection_date)

dummy_date <- min(nodes_df$infection_date, na.rm = TRUE) - 1
nodes_df$infection_date[is.na(nodes_df$infection_date)] <- dummy_date

expanded_dates <- seq(dummy_date, max(nodes_df$infection_date), by="day")
animation_data <- expand.grid(name = nodes_df$name, frame_time = expanded_dates) %>%
  left_join(nodes_df, by = "name") %>%
  mutate(visible = if_else(frame_time >= infection_date, 1, 0))

p <- ggraph(g, layout = 'tree') + 
  geom_edge_link(start_cap = circle(2, 'mm'), end_cap = circle(2, 'mm'),
                 color = 'gray20', alpha=0.9, 
                 arrow = arrow(type = "closed", length = unit(2, "mm"))) +
  geom_node_point(aes(color = status, shape = Serotype, alpha = visible), size = 5) +
  scale_color_manual(values = c("Subclinical" = "orange", "Clinical" = "steelblue", "Unknown" = "gray50")) +
  scale_alpha(range = c(0.1, 1)) +  
  theme_void() +
  labs(title = "Network Graph",
       subtitle = 'Date: {frame_time}') +
  transition_time(frame_time) +
  shadow_mark(alpha = 0.3, size = 3) 

fps_value <- 5 
animation_duration <- 20 #
pause_frames <- 50 

anim <- animate(p, width = 800, height = 600, fps = fps_value, duration = animation_duration, end_pause = pause_frames, renderer = gifski_renderer())
```

```{r eval=FALSE, echo=FALSE}
library(igraph)
library(ggraph)
library(ggplot2)
library(gganimate)
library(dplyr)
edges <- data.frame(from = all_networks$infector_ani, to = all_networks$infected_ani)

g <- graph_from_data_frame(edges, directed = TRUE)

V(g)$status <- "Unknown"

V(g)$status[V(g)$name %in% all_networks$infected_ani] <- all_networks$status[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$Serotype <- NA
V(g)$Serotype[V(g)$name %in% all_networks$infected_ani] <- 
  all_networks$Serotype[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$infection_date <- NA
V(g)$infection_date[V(g)$name %in% all_networks$infected_ani] <- 
  all_networks$infection_date[match(V(g)$name[V(g)$name %in% all_networks$infected_ani], all_networks$infected_ani)]

V(g)$infection_date <- as.Date(V(g)$infection_date)

nodes_df <- data.frame(name = V(g)$name, 
                       status = V(g)$status, 
                       Serotype = V(g)$Serotype, 
                       infection_date = V(g)$infection_date)

earliest_date <- min(nodes_df$infection_date, na.rm = TRUE)
dummy_date <- earliest_date - 1
nodes_df$infection_date[is.na(nodes_df$infection_date)] <- dummy_date

expanded_dates <- seq(dummy_date, max(nodes_df$infection_date), by = "day")
animation_data <- expand.grid(name = nodes_df$name, frame_time = expanded_dates) %>%
  left_join(nodes_df, by = "name") %>%
  mutate(visible = if_else(frame_time >= infection_date, 1, 0))

p <- ggraph(g, layout = 'tree') + 
  geom_edge_link(start_cap = circle(2, 'mm'), end_cap = circle(2, 'mm'),
                 color = 'gray20', alpha=0.9, 
                 arrow = arrow(type = "closed", length = unit(2, "mm"))) +
  geom_node_point(aes(color = status, shape = Serotype, filter = visible == 1), size = 5) +
  scale_color_manual(values = c("Subclinical" = "orange", "Clinical" = "steelblue", "Unknown" = "gray50")) +
  theme_void() +
  labs(title = "Network Graph",
       subtitle = 'Date: {frame_time}') +
  transition_time(frame_time) +
  shadow_mark(alpha = 0.3, size = 3) 


fps_value <- 5 
animation_duration <- 20 
pause_frames <- 50 

anim <- animate(p, width = 800, height = 600, fps = fps_value, duration = animation_duration, end_pause = pause_frames, renderer = gifski_renderer())
anim_save("network_animation_with_shadow.gif", anim)
```
