---
title: "Time Calibrated Phylogenies"
description: "Construction and evaluation of time calibrated trees for each serotype"
format:
  html: 
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options: 
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
knit_print.gt <- function(x, ...) {
  stringr::str_c(
    "<div style='all:initial';>\n", 
    gt::as_raw_html(x), 
    "\n</div>"
  ) |> 
    knitr::asis_output()
    
}
registerS3method(
  "knit_print", 'gt_tbl', knit_print.gt, 
  envir = asNamespace("gt") 
)
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(ape)
library(ggtree)
library(ggmap)
library(ggspatial)
library(dendextend)
library(pals)
library(gt)

## Custom Functions
source(here("R/utilities.R"))
source_dir(here("R"))

```


## Serotype A   
Check tree model statistics the *get_tracer_stats()* function calculates similar summary statistics as the Tracer software typically used with BEAST.  
```{r message=FALSE, warning=FALSE}
check_stats <- get_tracer_stats(here("local/beast/a_1/sero_a.log.txt"))

keep_stats <- c("joint", "prior", "likelihood", "treeModel.rootHeight", "age.root.",
                "treeLength", "tmrca.fmd_a_rev.", "clock.rate")

check_stats %>%
  filter(Parameter %in% keep_stats) %>%
  gt() %>%
  tab_header(
    title = md("Serotype A Stats")) %>%
  cols_width(starts_with("Parameter") ~ px(200),
             starts_with("label") ~ px(60),
             everything() ~ px(95)) %>%
  tab_options(table.font.size = "small",
              row_group.font.size = "small",
              stub.font.size = "small",
              column_labels.font.size = "medium",
              heading.title.font.size = "large",
              data_row.padding = px(2),
              heading.title.font.weight = "bold",
              column_labels.font.weight = "bold") %>%
  opt_stylize(style = 6, color = 'gray')
```

Load MCC tree
```{r fig.width=10, fig.height=18}
sero_A.tree <- read.nexus(here("local/beast/a_1/sero_a.mcc.tre"))

plot_time_tree(sero_A.tree, check_stats)

```

FMDV Effective Population Size    
Trees were run using several different clock and prior choices, all showed flat-line Ne.  The one here was the simplest, coalescent with contant population size.    
```{r fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
phylodynamic_process(sero_A.tree, check_stats, x_limits = c("2011-01-01", "2013-01-01"))
```


## Serotype Asia1   
```{r}
check_stats <- get_tracer_stats(here("local/beast/asia1_1/sero_asia1.log.txt"))

keep_stats <- c("joint", "prior", "likelihood", "treeModel.rootHeight", "age.root.",
                "treeLength", "tmrca.fmd_a_rev.", "clock.rate")

check_stats %>%
  filter(Parameter %in% keep_stats) %>%
  gt() %>%
  tab_header(
    title = md("Serotype Asia1 Stats")) %>%
  cols_width(starts_with("Parameter") ~ px(200),
             starts_with("label") ~ px(60),
             everything() ~ px(95)) %>%
  tab_options(table.font.size = "small",
              row_group.font.size = "small",
              stub.font.size = "small",
              column_labels.font.size = "medium",
              heading.title.font.size = "large",
              data_row.padding = px(2),
              heading.title.font.weight = "bold",
              column_labels.font.weight = "bold") %>%
  opt_stylize(style = 6, color = 'gray')
```

Load MCC tree
```{r fig.width=10, fig.height=18}
sero_Asia1.tree <- read.nexus(here("local/beast/asia1_1/sero_asia1.mcc.tre")) 
plot_time_tree(sero_Asia1.tree, check_stats, legend_pos = c(0.2, 0.5))

```

FMDV Effective Population Size    
Flat Ne, like the other serotypes  
```{r fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
phylodynamic_process(sero_Asia1.tree, check_stats, x_limits = c("2011-01-01", "2013-01-01"))
```

## Serotype O   
```{r}
check_stats <- get_tracer_stats(here("local/beast/o_1/sero_o.log.txt"))

keep_stats <- c("joint", "prior", "likelihood", "treeModel.rootHeight", "age.root.",
                "treeLength", "tmrca.fmd_a_rev.", "clock.rate")

check_stats %>%
  filter(Parameter %in% keep_stats) %>%
  gt() %>%
  tab_header(
    title = md("Serotype O Stats")) %>%
  cols_width(starts_with("Parameter") ~ px(200),
             starts_with("label") ~ px(60),
             everything() ~ px(95)) %>%
  tab_options(table.font.size = "small",
              row_group.font.size = "small",
              stub.font.size = "small",
              column_labels.font.size = "medium",
              heading.title.font.size = "large",
              data_row.padding = px(2),
              heading.title.font.weight = "bold",
              column_labels.font.weight = "bold") %>%
  opt_stylize(style = 6, color = 'gray')
```

Load MCC tree   
```{r fig.width=10, fig.height=18}
sero_O.tree <- read.nexus(here("local/beast/o_1/sero_o.mcc.tre"))

plot_time_tree(sero_O.tree, check_stats, legend_pos = c(0.2, 0.5))

```

FMDV Effective Population Size  
Flat Ne, like the other serotypes  
```{r fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
phylodynamic_process(sero_O.tree, check_stats, x_limits = c("2011-01-01", "2013-01-01"))
```







